### 2022-05-17 README.Rmd
Initial commit

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = FALSE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

# Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany

This repository contains the documentation, results, and code of a project evaluating the use of a semi-parametric nowcasting approach for COVID-19 hospitalisations in Germany. See the documentation for further details. This project is part of a [wider collaboration](https://covid19nowcasthub.de) assessing a range of nowcasting methods whilst providing an ensemble nowcast of COVID-19 Hospital admissions in Germany by date of positive test. This ensemble should be used for any policy related work rather than the nowcasts provided in this repository. See [here](https://covid19nowcasthub.de) for more on this nowcasting collaboration.

## Citation

If making use of the results of this analysis or reusing the analysis pipeline please cite:

If making using of the models evaluated in this analysis please also cite [`epinowcast`](https://epiforecasts.io/epinowcast):

> Sam Abbott (2021). epinowcast: Hierarchical nowcasting of right censored epidemological counts, DOI: 10.5281/zenodo.5637165

A BibTeX entry for LaTeX users is also available:

  @Article{,
    title = {epinowcast: Hierarchical nowcasting of right censored epidemological counts},
    author = {Sam Abbott},
    journal = {Zenodo},
    year = {2021},
    doi = {10.5281/zenodo.5637165},
  }

## Documentation

```{r, child = "docs/chunks/documentation.Rmd", eval = TRUE}
```

## Key files and folders

Folder/File | Purpose
---|---
[`writeup`](writeup/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
[`_targets.Rmd`](_targets.Rmd) | Analysis workflow for interactive use.
[`R`](R/) | R functions used in the analysis and for evaluation.
[`data`](data/) | Input data and summarised output generated by steps in the analysis.
[`analyses`](analyses/) | Ad-hoc analyses not part of the overarching workflow. This includes a synthetic case study and a simplified example using Germany hospitalisation data.
[`.devcontainer`](.devcontainer/) | Resources for reproducibility using `vscode` and `docker`.

## Dependencies

All dependencies can be installed using the following,

```{r}
remotes::install_dev_deps()
```

Alternatively a docker [container](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/.devcontainer/Dockerfile) and [image](https://github.com/epiforecasts/eval-germany-sp-nowcasting/pkgs/container/eval-germany-sp-nowcasting) is provided. An easy way to make use of this is using the Remote development extension of `vscode`.

## Analyses

This analysis in this repository has been implemented using the [`targets`](https://docs.ropensci.org/targets/) package and associated packages. The workflow is defined in [`_targets.md`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.md) and can be explored interactively using [`_targets.Rmd`](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.Rmd) `Rmarkdown` document. The workflow can be visualised as the following graph.

![](figures/targets-graph.png)

This complete analysis can be recreated using the following (note this may take quite some time even with a fairly large amount of available compute),

```{bash}
bash bin/update-targets.sh
```

Alternative the following `targets` functions may be used to interactively explore the workflow:

- Run the workflow sequentially.

```{r}
targets::tar_make()
```

- Run the workflow using all available workers.

```{r}
targets::tar_make_future(workers = future::availableCores())
```

- Explore a graph of the workflow.

```{r}
targets::tar_visnetwork(targets_only = TRUE)
```

Watch the workflow as it runs in a `shiny` app.

```{r}
targets::tar_watch(targets_only = TRUE)
```

To use our archived version of the interim results (and so avoid long run times) use the following to download it. Note that this process has not been rigorously tested across environments and so may not work seamlessly).

```{r}
source(here::here("R", "targetss-archive.R"))
get_targets_archive()
```