### 2022-05-17 _targets.Rmd

---
title: "Analysis pipeline: Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
author: "Sam Abbott"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Pipeline

The analysis pipeline for this work can be regenerated by rendering this file,

```{r, eval = FALSE}
rmarkdown::render("_targets.Rmd")
```

The pipeline can then be run using,

```{r, eval = FALSE}
tar_make()
```

The complete pipeline can be visualised using,

```{r, eval = FALSE}
tar_visnetwork()
```

Alternatively the pipeline can be explored interactively using this notebook or updated programmatically using the scripts in `bin`. We also provide an archived version of our `targets` workflow if only wanting to reproduce sections of our analysis. This can be downloaded using the following,

```{r, eval = FALSE}
source(here::here("R", "targets-archive.R"))
get_targets_archive()
```

# Setup

Set up the workflow pipeline and options. We first load the `targets` package and remove the potentially outdated workflow.

```{r}
library(targets)
library(tarchetypes)
library(data.table)
library(epinowcast)
library(ggplot2)
library(purrr, quietly = TRUE)
library(here)
library(future)
library(future.callr)
tar_unscript()
```

We now define shared global options across our workflow and load R functions from the `R` folder.

```{targets globals, tar_globals = TRUE}
library(targets)
library(tarchetypes)
library(cmdstanr)
library(data.table)
library(epinowcast)
library(ggplot2)
library(purrr, quietly = TRUE)
library(here)
library(future)
library(future.callr)
plan(callr)
functions <- list.files(here("R"), full.names = TRUE)
walk(functions, source)
rm("functions")
set_cmdstan_path()  
tar_option_set(
  packages = c("data.table", "epinowcast", "scoringutils", "ggplot2", "purrr",
               "cmdstanr", "here"),
  deployment = "worker",
  memory = "transient",
  workspace_on_error = TRUE,
  error = "continue",
  garbage_collection = TRUE
)
```

# Ingest data and stratify by location and report date

* Watch the URL at which hospitalisation data is hosted in order to trigger an update to the workflow when new data is uploaded

```{targets hospitalisation_url}
tar_url(
  hospitalisation_url,
  "https://raw.githubusercontent.com/KITmetricslab/hospitalization-nowcast-hub/main/data-truth/COVID-19/COVID-19_hospitalizations_preprocessed.csv" # nolint
)
```

* Download and process hospitalisation data. Also download and combined data on public holidays in Germany.

```{targets hospitalisations, tar_simple = TRUE}
get_germany_hospitalisations(url = hospitalisation_url)
```

* Define date to start nowcasting

```{targets start_date, tar_simple = TRUE}
as.Date("2021-10-01")
```

* Define dates to nowcast.

```{targets nowcast_dates, tar_simple = TRUE}
unique(
  hospitalisations[reference_date >= start_date]$reference_date
)
```

* Define age groups

```{targets age_groups, tar_simple = TRUE}
unique(hospitalisations$age_group)
```

* Define locations to nowcast using age group data (currently limited to just national level).

```{targets locations, tar_simple = TRUE}
"DE"
```

* Define locations in which to nowcast only for the overall count and not for each age group.

```{targets other_locations, tar_simple = TRUE}
setdiff(unique(hospitalisations$location), locations)
```

* Define maximum allowed reporting delay (in days).

```{targets max_report_delay, tar_simple = TRUE}
40
```

* Define hospitalisations by date of report. Note that here the target is defined to never update if data exists for that nowcast date. This is because the truth date for previous days can change slightly over time as negative cases are adjusted for leading to spurious refits.

```{targets hospitalisations_by_date_report}
tar_target(
  hospitalisations_by_date_report,
  enw_retrospective_data(
    hospitalisations,
    rep_date = nowcast_dates,
    ref_days = max_report_delay
  )[, nowcast_date := nowcast_dates],
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Define latest available data.

```{targets latest_hospitalisations, tar_simple = TRUE}
 enw_latest_data(hospitalisations)
```

* Make latest observations into 7 day incidence.

```{targets, latest_7day_hospitalisations, tar_simple = TRUE}
copy(latest_hospitalisations)[,
  confirm := frollsum(confirm, n =  7), by = c("age_group", "location")
][!is.na(confirm)]
```

* Define "complete" data (i.e more than 28 days have passed since first reported)

```{targets, complete_hospitalisations, tar_simple = TRUE}
latest_hospitalisations[reference_date < (max(reference_date) - 28)]
```

* Similarly, define "complete" 7 day hospitalisation incidence.

```{targets, complete_7day_hospitalisations, tar_simple = TRUE}
latest_7day_hospitalisations[reference_date < (max(reference_date) - 28)]
```

* Plot reporting delay percentage by date, location, and age group.

# Fit models and produce nowcasts

## Model and model settings

* Compile the model for multithread use.

```{targets, epinowcast_model}
tar_file(
  epinowcast_model,
  compile_model(),
  deployment = "main"
)
```
* Define stan settings shared across models and used for fitting

```{targets epinowcast_settings, tar_simple = TRUE}
list(
  save_warmup = FALSE,
  output_loglik = FALSE,
  pp = FALSE,
  iter_warmup = 1000,
  iter_sampling = 1000,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 1,
  adapt_delta = 0.95,
  show_messages = FALSE,
  refresh = 0
)
```

## Priors

* Define a set of uninformed priors using the defaults from `epinowcast`.

```{targets uninformed_priors, tar_simple = TRUE}
enw_priors()
```

* Define a set of observations to use as a source for informed priors. Here we use overall national hospitalisations and data from the 1st of July 2021.

```{targets prior_obs}
tar_target(
  prior_obs,
  enw_retrospective_data(
    hospitalisations[location == "DE"][age_group == "00+"],
    rep_date = as.Date("2021-07-01"), ref_days = max_report_delay
  ),
  cue = tar_cue(mode = "never")
)
```

* Fit a model to national overall hospitalisations only and extract posterior estimates for the delay distribution and overdispersion to use as informed priors for other models by assuming a normal distribution with their posterior standard deviations scaled by 10 times.

```{targets priors, tar_simple = TRUE} 
do.call(prior_epinowcast, c(
  list(
    prior_obs, max_delay = max_report_delay, scale = 10,
    priors = uninformed_priors
  ),
  epinowcast_settings
))
```

## Models

* Intercept only model.

```{targets fixed_nowcast}
tar_target(
  fixed_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = fixed_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Intercept model with day of the week reporting effects.

```{targets dow_nowcast}
tar_target(
  dow_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with day of the week effect reporting effects.

```{targets age_nowcast}
tar_target(
  age_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = age_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with a weekly random walk and a day of the week reporting effect.

```{targets week_nowcast}
tar_target(
  week_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = week_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Age group random effect model with an age group specific weekly random walk,  and a day of the week reporting effect.

```{targets age_week_nowcast}
tar_target(
  age_week_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report,
    tar_loc = locations,
    model = age_week_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations),
  cue = tar_cue(mode = "never")
)
```

* Independent age group model with a weekly random walk and a day of the week reporting model.

```{targets independent_nowcast}
tar_target(
  independent_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == age_groups],
    tar_loc = locations,
    model = independent_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations, age_groups),
  cue = tar_cue(mode = "never")
)
```

* Independent age group model with a weekly random walk and a day of the week reference and reporting model.

```{targets independent_ref_dow_nowcast}
tar_target(
  independent_ref_dow_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == age_groups],
    tar_loc = locations,
    model = independent_ref_dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, locations, age_groups),
  cue = tar_cue(mode = "never")
)
```

* Model for locations without age groups. As for the previous model this has a weekly random walk and a day of the week reporting and reference model. This was updated on the 6th of December 2021 from the independent model without a day of the week effect for reference date (to enable this note that we have overridden the default `cue` settings here so that once a model has been fit it is never run again for that combination of locations and dates regardless of upstream changes).

```{targets overall_only_nowcast}
tar_target(
  overall_only_nowcast,
  nowcast(
    obs = hospitalisations_by_date_report[age_group == "00+"],
    tar_loc = other_locations,
    model = independent_ref_dow_epinowcast,
    priors = priors,
    max_delay = max_report_delay,
    settings = epinowcast_settings
  ),
  cross(hospitalisations_by_date_report, other_locations),
  cue = tar_cue(mode = "never")
)
```

# Postprocess

* Combine nowcasts from each model

```{targets combined_nowcasts, tar_simple = TRUE}
rbindlist(list(
  fixed_nowcast,
  dow_nowcast,
  age_nowcast,
  week_nowcast,
  age_week_nowcast,
  independent_nowcast,
  overall_only_nowcast,
  independent_ref_dow_nowcast
), use.names = TRUE, fill = TRUE)[,
   model := factor(
    model,
    levels = c("Reference: Fixed, Report: Fixed",
               "Reference: Fixed, Report: Day of week",
               "Reference: Age, Report: Day of week",
               "Reference: Age and week, Report: Day of week",
               "Reference: Age and week by age, Report: Day of week",
               "Independent by age, Reference: Week, Report: Day of week",
               "Independent by age, Reference: Week and day of week, Report: Day of week")
   )
  ][, id := 1:.N]
```

* Extract summarised daily nowcast. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets summarised_nowcast, tar_simple = TRUE}
combined_nowcasts |> 
  adjust_posteriors(
    target = "daily", 
    max_ratio = 0.25, 
    rhat_bound = 1.1,
    per_dt_bound = 0.2
) |> 
  unnest_nowcasts(target = "daily") 
```

* Extract summarised 7 day nowcast. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets summarised_7day_nowcast, tar_simple = TRUE}
combined_nowcasts |> 
  adjust_posteriors(
    target = "seven_day", 
    max_ratio = 0.25, 
    rhat_bound = 1.1,
    per_dt_bound = 0.2
) |> 
  unnest_nowcasts(target = "seven_day") 
```

* Save daily nowcasts stratified by nowcasting date.

```{targets save_daily_nowcasts}
tar_file(
  save_daily_nowcasts,
  summarised_nowcast[nowcast_date == nowcast_dates] |>
    save_csv(
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/daily"),
      allow_empty = FALSE
    ),
  map(nowcast_dates),
  cue = tar_cue(mode = "never")
)
```

* Save 7 day nowcasts stratified by nowcasting date

```{targets save_7day_nowcasts}
tar_file(
  save_7day_nowcasts,
  summarised_7day_nowcast[nowcast_date == nowcast_dates] |>
    save_csv(
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/seven_day"),
      allow_empty = FALSE
    ),
  map(nowcast_dates),
  cue = tar_cue(mode = "never")
)
```

* Define and format the hierarchical nowcast for submission. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values).

```{targets hierarchical_submission_nowcast}
tar_target(
  hierarchical_submission_nowcast,
  age_week_nowcast[nowcast_date == nowcast_dates] |>
    adjust_posteriors(
      target = "seven_day", 
      max_ratio = 0.25, 
      rhat_bound = 1.1,
      per_dt_bound = 0.2
    ) |>
    select_var("seven_day") |>
    rbindlist() |>
    format_for_submission(),
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Format and save 7 day hierarchical nowcasts for hub submission.

```{targets save_hierarchical_submission}
tar_file(
  save_hierarchical_submission,
  save_csv(
    hierarchical_submission_nowcast,
      filename = paste0(nowcast_dates, ".csv"),
      path = here("data/nowcasts/submission/hierarchical")
  ),
  map(hierarchical_submission_nowcast, nowcast_dates)
)
```

* Define and format the independent nowcast for submission. As a temporary measure here we adjust quantiles that are more than 25% of the median when there is evidence of a fitting issue (based on divergent transistions and R hat values). On the 6th of December the submission model was updated to include a reference day of the week effect.

```{targets independent_submission_nowcast}
tar_target(
  independent_submission_nowcast,
  rbind(
    independent_ref_dow_nowcast[nowcast_date == nowcast_dates],
    overall_only_nowcast[nowcast_date == nowcast_dates]
  ) |> 
   adjust_posteriors(
      target = "seven_day", 
      max_ratio = 0.25, 
      rhat_bound = 1.1,
      per_dt_bound = 0.2
    ) |>
    select_var("seven_day") |>
    rbindlist() |>
    format_for_submission(),
  map(nowcast_dates),
  iteration = "list",
  cue = tar_cue(mode = "never")
)
```

* Format and save 7 day independent nowcasts for hub submission.

```{targets save_independent_submission}
tar_file(
  save_independent_submission,
  save_csv(
    independent_submission_nowcast,
    filename = paste0(nowcast_dates, ".csv"),
    path = here("data/nowcasts/submission/independent")
  ),
  map(independent_submission_nowcast, nowcast_dates)
)
```

* Save latest hospitalisation data

```{targets save_latest_daily_hospitalisations}
tar_file(
  save_latest_daily_hospitalisations,
  save_csv(
    latest_hospitalisations,
    filename = paste0("daily.csv"),
    path = here("data/observations")
  )
)
```

* Save 7 day hospitalisation data

```{targets save_latest_7day_hospitalisations}
tar_file(
  save_latest_7day_hospitalisations,
  save_csv(
    latest_7day_hospitalisations,
    filename = paste0("seven_day.csv"),
    path = here("data/observations")
  )
)
```

# Evaluation

* Extract and save model fitting diagnostics.

```{targets diagnostics, tar_simple = TRUE}
combined_nowcasts[, c("daily", "seven_day") := NULL]
```

```{targets save_diagnostics}
list(
  tar_file(
    save_all_diagnostics,
      save_csv(
        diagnostics,
        filename = "all.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_high_rhat_diagnostics,
      save_csv(
        diagnostics[max_rhat > 1.05],
        filename = "high-rhat.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_high_divergent_transitions,
      save_csv(
        diagnostics[per_divergent_transitions > 0.1],
        filename = "high-divergent-transitions.csv",
        path = here("data/diagnostics")
      )
  ),
  tar_file(
    save_failures,
      save_csv(
        diagnostics[failure == TRUE],
        filename = "fitting-failed.csv",
        path = here("data/diagnostics")
      )
  )
)
```

* Extract and save model run-time at the national level aggregated across age-group.

```{targets save-run-time}
tar_file(
  save_run_time,
  combined_nowcasts |>
    summarise_runtimes() |>
    save_csv(
      filename = "run-times.csv",
      path = here("data/diagnostics")
    )
)
```

* Filter nowcasts to only include those with "complete" (more than 28 days of reports) data for all age groupsand with horizons between 0 days and -7 days from the nowcast date.

```{targets scored_nowcasts, tar_simple = TRUE}
summarised_nowcast[location == locations][
  reference_date < (max(nowcast_date) - 28)][,
  holiday := NULL][,
  horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
  horizon >= -7
]
```

* Score daily nowcasts overall, by age group, by horizon, by nowcast date, and by reference date on both the natural and log scales (corresponding to absolute and relative scoring). These summarised scores are then saved to `data/scores`.

```{targets, score}
tar_map(
  list(score_by = list(
    "overall", "age_group", "horizon", "reference_date", "nowcast_date"
  )),
  tar_target(
    scores,
    enw_score_nowcast(
      scored_nowcasts, complete_hospitalisations, 
      summarise_by = drop_string(c(score_by, "model"), "overall"),
      log = FALSE
    )
  ),
  tar_target(
    log_scores,
    enw_score_nowcast(
      scored_nowcasts, complete_hospitalisations, 
      summarise_by = drop_string(c(score_by, "model"), "overall"),
      log = TRUE
    )
  ),
  tar_file(
    save_scores,
    save_csv(
      rbind(scores[, scale := "natural"], log_scores[, scale := "log"]),
      filename = paste0(paste(score_by, sep = "-"), ".csv"),
      path = here("data/scores")
    )
  )
)
```

# Visualise

**Currently these are not produced. See the real time report instead**

* Plot most recent daily nowcast by location, age group, and model.

```{r plot-latest-nowcast, eval = FALSE}
tar_target(
  plot_latest_nowcast,
  enw_plot_nowcast_quantiles(
    summarised_nowcast[nowcast_date == max(nowcast_date)][
                       location == locations][
                       reference_date >= (nowcast_date - 28)]
  ) +
  facet_grid(vars(age_group), vars(model), scales = "free_y"),
  map(locations),
  iteration = "list"
)
```

* Plot most recent seven day nowcast by location, age group, and model.

```{r plot-latest-7day-nowcast, eval = FALSE}
tar_target(
  plot_latest_7day_nowcast,
  enw_plot_nowcast_quantiles(
    summarised_7day_nowcast[nowcast_date == max(nowcast_date)][
                            location == locations][
                            reference_date >= (nowcast_date - 28)]
  ) +
  facet_grid(vars(age_group), vars(model), scales = "free_y"),
  map(locations),
  iteration = "list"
)
```

* Plot daily nowcasts at each horizon from the date of the nowcast to 7 days previously by location, age group and model.

```{r plot_nowcast_horizon, eval = FALSE}
tar_map(
  list(horizons = 0:7),
  tar_target(
    plot_nowcast_horizon,
    enw_plot_nowcast_quantiles(
      scored_nowcasts[location == locations][horizon == -horizons][,
                      confirm := NA],
      latest_obs = latest_hospitalisations[
        location == locations][
        reference_date >= min(scored_nowcasts$reference_date)][
        reference_date <= max(scored_nowcasts$reference_date)
      ],
      log = TRUE
    ) +
    facet_grid(vars(age_group), vars(model), scales = "free_y"),
    map(locations),
    iteration = "list"
  )
)
```

* Plot 7 day nowcasts at each horizon from the date of the nowcast to 7 days previously by location, age group and model.

```{r plot_7day_nowcast_horizon, eval = FALSE}
tar_map(
  list(horizons = 0:7),
  tar_target(
    plot_7day_nowcast_horizon,
    enw_plot_nowcast_quantiles(
      summarised_7day_nowcast[
        reference_date < (max(nowcast_date) - 28)][,
        holiday := NULL][,
        horizon := as.numeric(as.Date(reference_date) - nowcast_date)][
        location == locations][
        horizon == -horizons][,
        confirm := NA],
      latest_obs = latest_7day_hospitalisations[
        location == locations][
        reference_date >= min(scored_nowcasts$reference_date)][
        reference_date <= max(scored_nowcasts$reference_date)
      ],
      log = TRUE
    ) +
    facet_grid(vars(age_group), vars(model), scales = "free_y"),
    map(locations),
    iteration = "list"
  )
)
```

* Plot daily nowcasts for locations without age groups stratified nowcasts.

* Plot 7 day nowcasts for locations without age group stratified nowcasts.

* Plot scores relative to the baseline by location and age group on both the natural and log scale.