### 2022-05-17 docs/chunks/documentation.Rmd

Document | Purpose
---|---
[Summary](https://epiforecasts.io/eval-germany-sp-nowcasting/) | A summary of this work.
[Paper](https://epiforecasts.io/eval-germany-sp-nowcasting/paper.pdf) | The academic paper write up of this work.
[Supplementary information](https://epiforecasts.io/eval-germany-sp-nowcasting/si.html) | The supplementary information for the write up of this work.
[Real-time model evaluation](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time/) | A report visualising and evaluating nowcasts from the various model configurations considered here in real-time.
[Real-time method evaluation](https://epiforecasts.io/eval-germany-sp-nowcasting/real-time-method-comparison/) | A report visualising and evaluating nowcasts from the various methods (from this project and other groups) submitted to the [Germany nowcasting hub](https://covid19nowcasthub.de) in real-time.
[Project README](https://github.com/epiforecasts/eval-germany-sp-nowcasting) | Overarching project README. Includes links to resources, a summary of key files, and reproducibility information.
[Analysis pipeline](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/_targets.md) | The `targets` based analysis pipeline.
[Analysis archive](https://github.com/epiforecasts/eval-germany-sp-nowcasting/releases/tag/latest) | An archived version of the `_targets` directory. Download using `get_targets_archive()`.
[Data](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/data/README.md) | Documentation for input data and summarised output from the analysis.
[bin](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/bin/README.md) | Documentation for orchestration of nowcast estimation, publishing, and archiving.
[News](https://github.com/epiforecasts/eval-germany-sp-nowcasting/blob/main/NEWS.md) | Dated development notes.
[epinowcast](https://epiforecasts.io/epinowcast/index.html) | The documentation for `epinowcast` the R package used to implement the models evaluated here. See this [case study](https://epiforecasts.io/epinowcast/articles/germany-age-stratified-nowcasting.html) for a simplified version of this analysis.
[Germany nowcasting hub](https://covid19nowcasthub.de)  | The homepage (containing a dashboard and information) for the Germany nowcasting hub project to which nowcasts from this evaluation are submitted along with others produced by independent groups.