### 2022-05-17 writeup/si.Rmd

---
title: "Evaluating Semi-Parametric Nowcasts of COVID-19 Hospital Admissions in Germany"
subtitle: "Supplementary Information"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    toc_depth: 4
    includes:
      before_body: ../docs/header.html
      after_body: ../docs/footer.html
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
bibliography: library.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa-numeric-superscript-brackets.csl
header-includes:
   - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE,
  fig.width = 9, fig.height = 9,
  dpi = 320,
  fig.path = "figures/"
)
```

## Authors

S. Abbott (1), S. Funk (1)

Correspondence to: sam.abbott@lshtm.ac.uk

## Affiliations

1. Center for the Mathematical Modelling of Infectious Diseases, London School of Hygiene & Tropical Medicine, London WC1E 7HT, United Kingdom

## Overview

This analysis uses the `targets` package for reproducibility. The full analysis structure can be seen below:

```{r, message = FALSE, eval = FALSE}
library(targets)
tar_glimpse()
```

![](../figures/targets-graph.png)

## References

<div id = 'refs'></div>