### 2022-08-23 README.Rmd
repository clean up

---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, eval = FALSE
)
```

# Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

This repository contains the documentation, results, and code of a project evaluating a simplified forecasting model in comparison the ECDC forecasting hub ensemble. See the documentation for further details.

## Citation

## Project structure

Folder | Purpose
---|---

[`ecdc-weekly-growth-forecasts`](ecdc-weekly-growth-forecasts/) | The code for the simplified forecasting model evaluated in this work.
[`data-raw`](data-raw/) | Raw input data and scripts required to download and process it.
[`data`](data/) | Processed data from `data-raw` ready to be used in the paper analysis.
[`R`](R/) | R functions used in the analysis and for evaluation.
[`paper`](paper/) | Summary paper and additional supplementary information as `Rmarkdown` documents.
[`.devcontainer`](.devcontainer/) | Resources for reproducibility using `vscode` and `docker`.

## Dependencies

Dependencies are managed using [`renv`].

Alternatively a docker [container](https://github.com/epiforecasts/simplfied-forecaster-evaluation/blob/main/.devcontainer/Dockerfile) and [image](https://github.com/epiforecasts/simplfied-forecaster-evaluation/pkgs/container/simplfied-forecaster-evaluation) is provided. An easy way to make use of this is using the Remote development extension of `vscode`.